Cascaded Approximation of Gaussian Blur

### 2015-05-22 article.tex
Minor fixes

\title{Cascaded Approximation of Gaussian Blur}
\author{Vabishchevich Nikolay}
\date{May 19, 2015}
\maketitle

@@

In this article I'm using two different sets of grid point positions. The first one is grid point at
every integer, $F^+_k$ (I call it even and mark with ``$+$''). The second one is grid points positioned
halfway between adjacent integers, $F^-_{k+1/2}$ (I call it odd and mark with ``$-$'').
Correspondingly there are two types of convolution filters: even
\begin{equation}\label{conv+}
    R^\pm_p = \sum_{k\in Z} F^\pm_{p+k}K^+_k,
\end{equation}
and odd
\begin{equation}\label{conv-}
    R^\mp_p = \sum_{k\in Z} F^\pm_{p+k+\frac12}K^-_{k+\frac12},
\end{equation}
where $K^\pm_p$ is filter kernel. Even filter doesn't change source oddness but odd filter do (hence
the name). Therefore our target problem is to efficiently calculate approximation of
\begin{equation}
    R^\pm_p = \sum_{k\in Z} F^\pm_{p+k}G_k,
\end{equation}
where $F^\pm_p$ is the source function and $G_k$ is even gaussian kernel,
\begin{equation}\label{gamma}
    G_k = \Gamma(\sigma^2;\, k) \equiv
        \frac1{\sqrt{2\pi\sigma^2}}\exp\left\{-\frac{k^2}{2\sigma^2}\right\},
\end{equation}
especially for large standard deviation $\sigma$.

@@

\pgfplotsset{spectrum/.style = {
    width = 6cm, height = 6cm, domain = -1:1, samples = 256,
    xtick = {-1, 0, 1}, xticklabels = {$-\pi$, $0$, $\pi$}, minor x tick num = 1,
    xmin = -1, xmax = 1, xminorgrids = true, xmajorgrids = true,
    ymin = 0, ymax = 1, yminorgrids = true, ymajorgrids = true,
}}
\begin{wrapfigure}{r}{0pt}
    \begin{tikzpicture}
        \begin{axis} [ spectrum, title = $G^+(\varphi)$ ]
            \addplot [ sharp plot, line width = 1pt, red ] gnuplot {
                exp(-8*(x*pi)^2)
            };
        \end{axis}
    \end{tikzpicture}
    \caption{Gaussian spectrum (\ref{gauss}) for $\sigma = 4$.}
\end{wrapfigure}

Going back to our gaussian filter (\ref{gamma}), its Fourier transform is another gaussian function
\begin{equation}\label{gauss}
    G(\varphi) \approx \exp\left\{-\frac{\sigma^2\varphi^2}2\right\}.
\end{equation}
That function has two different regions. The first is almost zero at high frequency and the second
is bell-like shape at low frequency. Correspondingly the task of approximation can be divided into
two subtasks described in the following sections.

@@

The only way to avoid large filter kernels in approximation of gaussian blur with large radius is to
somehow reduce resolution before filtering. That can be done through successive steps of shrinking by
factor of two, then applying main filter and then expanding twofold the same number of steps.
Central main filter will be examined in the next section and this section concentrates on scaling.

@@

For last task of approximating central gaussian shape I've decided to use generic 9-tap filter as
the main workhorse. Under that name I imply even symmetric conservative convolutional filter of the
following form
\begin{equation}\label{main}
    K_k = \left\{
        \begin{array}{ll}
            1 - \sum_{i=1}^4 C_i,& k = 0;\\
            \frac12 C_i,& k = m_i;\\
            0,& \text{otherwise};
        \end{array}
    \right.
\end{equation}
where $m_i$, $i=1, 2, 3, 4$, are sampling positions ($0 < m_1 < m_2 < m_3 < m_4$) and $C_i$ are
coefficients. Fourier transform of this filter kernel is
\begin{equation}
    K(\varphi) = 1 - \sum_{i=1}^4 C_i[1 - \cos(m_i\varphi)].
\end{equation}
Further research shows that this filter isn't enough as central filter so I combine it with
additional stage of even binomial filter $\hat B^{2l}$ ($B^0$ mean ``do nothing''). Therefore
complete approximation of gaussian blur take the form of
\begin{equation}
    \hat G_{\text{apx}} = \hat Q^n\hat K\hat B^{2l}\hat P^n.
\end{equation}
Disregarding frequency crosstalk, diagonal part of that operator in Fourier space is
\begin{equation}
    G_{\text{apx}}(\varphi) = \left\{1 - \sum_{i=1}^4 C_i[1 - \cos(2^nm_i\varphi)]\right\}
        \cos^{2l}(2^{n-1}\varphi) \prod_{k=0}^{n-1} \cos^{10}(2^{k-1}\varphi).
\end{equation}
Outside of the region $[-\pi/2^n, \pi/2^n]$ that should be already near zero, so we focus only on
center
\begin{equation}
    G_{\text{apx}}\left(\frac\varphi{2^n}\right) =
        \left\{1 - \sum_{i=1}^4 C_i[1 - \cos(m_i\varphi)]\right\}
        \cos^{2l}\frac\varphi2 \prod_{k=1}^n \cos^{10}\frac\varphi{2^{k+1}},
\end{equation}
where that expression should be near
\begin{equation}
    G\left(\frac\varphi{2^n}\right) = \exp\left\{-\frac{\sigma^2\varphi^2}{2\cdot4^n}\right\}.
\end{equation}

One of the possible ways to find good kernel coefficients $C_i$ is to solve least squares problem.
Denote
\begin{equation}
    p(\varphi) \equiv \cos^{2l}\frac\varphi2 \prod_{k=1}^n \cos^{10}\frac\varphi{2^{k+1}}
\end{equation}
and
\begin{equation}
    g(\varphi) \equiv \exp\left\{-\frac{\sigma^2\varphi^2}{2\cdot4^n}\right\},
\end{equation}
then the problem is to minimize
\begin{equation*}
    \lambda = \int_{-\pi}^\pi \left[G\left(\frac\varphi{2^n}\right) -
        G_{\text{apx}}\left(\frac\varphi{2^n}\right)\right]^2 w(\varphi) d\varphi =
\end{equation*}
\begin{equation*}
    \int_{-\pi}^\pi \left[g(\varphi) -
        \left\{1 - \sum_{i=1}^4 C_i[1 - \cos(m_i\varphi)]\right\}p(\varphi)\right]^2
        w(\varphi) d\varphi =
\end{equation*}
\begin{equation}\label{lambda}
    \int_{-\pi}^\pi \left[g(\varphi) - p(\varphi) +
        p(\varphi)\sum_{i=1}^4 C_i[1 - \cos(m_i\varphi)]\right]^2 w(\varphi) d\varphi,
\end{equation}
where $w(\varphi) \approx 1$ is additional weight. As usual, calculating partial derivatives
\begin{equation*}
    0 = \frac12\frac{\partial\lambda}{\partial C_k} =
    \int_{-\pi}^\pi \left[g(\varphi) - p(\varphi) +
        p(\varphi)\sum_{i=1}^4 C_i[1 - \cos(m_i\varphi)]\right]
        p(\varphi)w(\varphi) [1 - \cos(m_k\varphi)] d\varphi =
\end{equation*}
\begin{equation}
    \int_{-\pi}^\pi [g(\varphi) - p(\varphi)]p(\varphi)w(\varphi) [1 - \cos(m_k\varphi)] d\varphi +
    \sum_{i=1}^4 C_i\int_{-\pi}^\pi p^2(\varphi)w(\varphi)
        [1 - \cos(m_i\varphi)][1 - \cos(m_k\varphi)] d\varphi.
\end{equation}
After defining
\begin{equation}
    U_k \equiv \frac1{2\pi}\int_{-\pi}^\pi p^2(\varphi)w(\varphi)e^{-ip\varphi}d\varphi
\end{equation}
and
\begin{equation}
    V_k \equiv \frac1{2\pi}\int_{-\pi}^\pi g(\varphi)p(\varphi)w(\varphi)e^{-ip\varphi}d\varphi
\end{equation}
we can get
\begin{equation}
    0 = \frac1{2\pi}\frac{\partial\lambda}{\partial C_k} = 2[V_0 - V_{m_k} - U_0 + U_{m_k}] +
        \sum_{i=1}^4 C_i[2U_0 - 2U_{m_i} - 2U_{m_k} + U_{m_i+m_k} + U_{|m_i-m_k|}].
\end{equation}
Resulting symmetric system of four linear equations
\begin{equation}
    \sum_{i=1}^4 C_i[2U_0 - 2U_{m_i} - 2U_{m_k} + U_{m_i+m_k} + U_{|m_i-m_k|}] =
        2[U_0 - U_{m_k} - V_0 + V_{m_k}],
\end{equation}
can be solved through standard methods of linear algebra.

The real problem here is to efficiently calculate $U_k$ and $V_k$. One method is to use crude
approximation
\begin{equation}
    p(\varphi) \approx \exp\left\{-\frac{\alpha\varphi^2}2\right\},
\end{equation}
where
\begin{equation}\label{alpha}
    \alpha = \frac l2 + \frac56(1 - 4^{-n}).
\end{equation}
In that case
\begin{equation}
    U_k \approx \frac1{2\pi}\int_{-\pi}^\pi
        \exp\left\{-\alpha\varphi^2\right\}e^{-ip\varphi}d\varphi \approx \Gamma(2\alpha;\, k),
\end{equation}
\begin{equation}\label{V-gauss}
    V_k \approx \frac1{2\pi}\int_{-\pi}^\pi
        \exp\left\{-\left[\frac{\sigma^2}{4^n} + \alpha\right]\frac{\varphi^2}2\right\}
        e^{-ip\varphi}d\varphi \approx \Gamma(\sigma^24^{-n} + \alpha;\, k),
\end{equation}
where $\Gamma$ is normalized gaussian function defined in (\ref{gamma}).

Much better method is to choose $w(\varphi)$ such that
\begin{equation}
    p(\varphi)w(\varphi) = \exp\left\{-\frac{\alpha\varphi^2}2\right\}
\end{equation}
holds exactly. In that case (\ref{V-gauss}) holds exactly too and for $U_k$ we
have
\begin{equation}
    U_k = \frac1{2\pi}\int_{-\pi}^\pi \exp\left\{-\frac{\alpha\varphi^2}2\right\}
        p(\varphi)e^{-ip\varphi}d\varphi =
    \frac1{2\pi}\int_{-\pi}^\pi \exp\left\{-\frac{\alpha\varphi^2}2\right\}
        B^{2l}(\varphi)D(\varphi)e^{-ip\varphi}d\varphi,
\end{equation}
where $D$ is effective rescale filter kernel with
\begin{equation}
    D(\varphi) \equiv \prod_{k=1}^n \cos^{10}\frac\varphi{2^{k+1}}.
\end{equation}
Therefore $U_k$ can be calculated in the following way: construct gaussian function
$\Gamma(\alpha;\, k)$, then apply filter $\hat D$, than apply binomial filter of order $2l$.
Alternatively that can be written in operator notation as
\begin{equation}
    \hat U = \hat B^{2l}\hat D\,\hat\Gamma(\alpha;\, \cdot).
\end{equation}

@@

Third method is usually slightly better than second and much better then first, but precise methods
require good approximation of effective rescale filter $D$. Experiments shows that tight 7-tap
approximation
\begin{equation}
    D(\varphi) \approx D_0 + 2\sum_{k=1}^3 D_k\cos(k\varphi)
\end{equation}
is quite good. Coefficients here should be chosen in such a way that equation is exact up to 6
term of Taylor series around $\varphi = 0$. Corresponding problem can be solved analytically
and give the following results
\newcommand\0{\phantom{0}}
\begin{align}
    D_0 &= (   5204 +   2520\cdot4^{-n} +   1092\cdot16^{-n} +   3280\cdot64^{-n}) / 12096,\\
    D_1 &= (   2943 - \0 210\cdot4^{-n} - \0 273\cdot16^{-n} -   2460\cdot64^{-n}) / 12096,\\
    D_2 &= (\0  486 - \0 924\cdot4^{-n} - \0 546\cdot16^{-n} + \0 984\cdot64^{-n}) / 12096,\\
    D_3 &= (\0\0 17 - \0 126\cdot4^{-n} + \0 273\cdot16^{-n} - \0 164\cdot64^{-n}) / 12096.
\end{align}
Fig.~\ref{D} shows that such approximation noticeably differ from exact curve only at region near
$\varphi = \pm\pi$ where the target gaussian is near zero anyway. Accurate approximation of
effective rescale filter kernel concludes calculation of the main filter coefficients, and in the
following section I study the task of choosing remaining parametrs $n$, $l$ and $m_i$.

@@

Combined effect of all stages of gaussian blur approximation can be written in generic form
\begin{equation}
    R_i = \sum_{j\in Z} K_{ij}F_j,
\end{equation}
where
\begin{equation}
    K_{ij} \approx G_{i-j} = \Gamma(\sigma^2;\, i - j) =
        \frac1{\sqrt{2\pi\sigma^2}}\exp\left\{-\frac{(i - j)^2}{2\sigma^2}\right\}.
\end{equation}
Assuming that filter doesn't distort constant function (if $F_i = c$ then $R_i = c$) or
alternatively
\begin{equation}
    \sum_{j\in Z} K_{ij} = 1,
\end{equation}
and source values lie between $0$ and $1$,
\begin{equation}
    0 \le F_i \le 1,
\end{equation}
we can calculate exact upper bound of error $E^{\max}_i = \sup_F |E_i|$, where
\begin{equation}\label{E}
    E_i = \sum_{j\in Z} [K_{ij} - G_{i-j}]F_j.
\end{equation}
Due to zero error in case of zero source function
\begin{equation}
    E^{\max}_i = \sup_F |E_i| = \max\{-\inf_F E_i,\; \sup_F E_i\}.
\end{equation}
To calculate maximum positive and negative error one can choose $F_j$ in such a way that sum from
(\ref{E}) reduces to sum of only positive or negative terms,
\begin{align}
    \inf_F E_i &= \sum_{j\in Z, K_{ij} < G_{i-j}} [K_{ij} - G_{i-j}],\\
    \sup_F E_i &= \sum_{j\in Z, K_{ij} > G_{i-j}} [K_{ij} - G_{i-j}].
\end{align}
It's easy to see that
\begin{align}
    \sup_F E_i + \inf_F E_i &= \sum_{j\in Z} [K_{ij} - G_{i-j}] = 1 - 1 = 0,\\
    \sup_F E_i - \inf_F E_i &= \sum_{j\in Z} |K_{ij} - G_{i-j}|,
\end{align}
and
\begin{equation}
    E^{\max}_i = \max\{-\inf_F E_i,\; \sup_F E_i\} = \sup_F E_i =
        \frac12\sum_{j\in Z} |K_{ij} - G_{i-j}|.
\end{equation}
That's the final form of error estimation I'll use later.

>>>>

\title{Cascaded Approximation of Gaussian Blur}
\author{Vabishchevich Nikolay}
\date{May 22, 2015}
\maketitle

@@

In this article I'm using two different sets of grid point positions. The first one is grid point at
every integer, $F^+_k$ (I call it even and mark with ``$+$''). The second one is grid points positioned
halfway between adjacent integers, $F^-_{k+1/2}$ (I call it odd and mark with ``$-$'').
Correspondingly there are two types of convolution filters: even
\begin{equation}\label{conv+}
    R^\pm_p = \sum_{k\in Z} F^\pm_{p+k}K^+_k,
\end{equation}
and odd
\begin{equation}\label{conv-}
    R^\mp_p = \sum_{k\in Z} F^\pm_{p+k+\frac12}K^-_{k+\frac12},
\end{equation}
where $K^\pm_p$ is filter kernel. Even filter doesn't change source oddness but odd filter do (hence
the name). Therefore our target problem is to efficiently calculate approximation of
\begin{equation}
    R^\pm_p = \sum_{k\in Z} F^\pm_{p+k}G_k,
\end{equation}
where $F^\pm_p$ is the source function and $G_k$ is even Gaussian kernel,
\begin{equation}\label{gamma}
    G_k = \Gamma(\sigma^2;\, k) \equiv
        \frac1{\sqrt{2\pi\sigma^2}}\exp\left\{-\frac{k^2}{2\sigma^2}\right\},
\end{equation}
especially for large standard deviation $\sigma$.

@@

\pgfplotsset{spectrum/.style = {
    width = 6cm, height = 6cm, domain = -1:1, samples = 256,
    xtick = {-1, 0, 1}, xticklabels = {$-\pi$, $0$, $\pi$}, minor x tick num = 1,
    xmin = -1, xmax = 1, xminorgrids = true, xmajorgrids = true,
    ymin = 0, ymax = 1, yminorgrids = true, ymajorgrids = true,
}}
\begin{wrapfigure}{r}{0pt}
    \begin{tikzpicture}
        \begin{axis} [ spectrum, title = $G(\varphi)$ ]
            \addplot [ sharp plot, line width = 1pt, red ] gnuplot {
                exp(-8*(x*pi)^2)
            };
        \end{axis}
    \end{tikzpicture}
    \caption{Gaussian spectrum (\ref{gauss}) for $\sigma^2 = 16$.}
\end{wrapfigure}

Going back to our Gaussian filter (\ref{gamma}), its Fourier transform is another Gaussian function
\begin{equation}\label{gauss}
    G(\varphi) \approx \exp\left\{-\frac{\sigma^2\varphi^2}2\right\}.
\end{equation}
That function has two different regions. The first is almost zero at high frequency and the second
is bell-like shape at low frequency. Correspondingly the task of approximation can be divided into
two subtasks described in the following sections.

@@

The only way to avoid large filter kernels in approximation of Gaussian blur with large radius is to
somehow reduce resolution before filtering. That can be done through successive steps of shrinking by
factor of two, then applying main filter and then expanding twofold the same number of steps.
Central main filter will be examined in the next section and this section concentrates on scaling.

@@

For last task of approximating central Gaussian shape I've decided to use generic 9-tap filter as
the main workhorse. Under that name I imply even symmetric conservative convolutional filter of the
following form
\begin{equation}\label{main}
    K_k = \left\{
        \begin{array}{ll}
            1 - 2\sum_{i=1}^4 C_i,& k = 0;\\
            C_i,& k = m_i;\\
            0,& \text{otherwise};
        \end{array}
    \right.
\end{equation}
where $m_i$, $i=1, 2, 3, 4$, are sampling positions ($0 < m_1 < m_2 < m_3 < m_4$) and $C_i$ are
coefficients. Fourier transform of this filter kernel is
\begin{equation}
    K(\varphi) = 1 - \sum_{i=1}^4 C_i[2 - 2\cos(m_i\varphi)].
\end{equation}
Further research shows that this filter isn't enough as central filter so I combine it with
additional stage of even binomial filter $\hat B^{2l}$ ($B^0$ mean ``do nothing''). Therefore
complete approximation of Gaussian blur take the form of
\begin{equation}
    \hat G_{\text{apx}} = \hat Q^n\hat K\hat B^{2l}\hat P^n.
\end{equation}
Disregarding frequency crosstalk, diagonal part of that operator in Fourier space is
\begin{equation}
    G_{\text{apx}}(\varphi) = \left\{1 - \sum_{i=1}^4 C_i[2 - 2\cos(2^nm_i\varphi)]\right\}
        \cos^{2l}(2^{n-1}\varphi) \prod_{k=0}^{n-1} \cos^{10}(2^{k-1}\varphi).
\end{equation}
Outside of the region $[-\pi/2^n, \pi/2^n]$ that should be already near zero, so we focus only on
center
\begin{equation}
    G_{\text{apx}}\left(\frac\varphi{2^n}\right) =
        \left\{1 - \sum_{i=1}^4 C_i[2 - 2\cos(m_i\varphi)]\right\}
        \cos^{2l}\frac\varphi2 \prod_{k=1}^n \cos^{10}\frac\varphi{2^{k+1}},
\end{equation}
where that expression should be near
\begin{equation}
    G\left(\frac\varphi{2^n}\right) = \exp\left\{-\frac{\sigma^2\varphi^2}{2\cdot4^n}\right\}.
\end{equation}

One of the possible ways to find good kernel coefficients $C_i$ is to solve least squares problem.
Denote
\begin{equation}
    p(\varphi) \equiv \cos^{2l}\frac\varphi2 \prod_{k=1}^n \cos^{10}\frac\varphi{2^{k+1}}
\end{equation}
and
\begin{equation}
    g(\varphi) \equiv \exp\left\{-\frac{\sigma^2\varphi^2}{2\cdot4^n}\right\},
\end{equation}
then the problem is to minimize
\begin{equation*}
    \lambda = \int_{-\pi}^\pi \left[G\left(\frac\varphi{2^n}\right) -
        G_{\text{apx}}\left(\frac\varphi{2^n}\right)\right]^2 w(\varphi) d\varphi =
\end{equation*}
\begin{equation*}
    \int_{-\pi}^\pi \left[g(\varphi) -
        \left\{1 - \sum_{i=1}^4 C_i[2 - 2\cos(m_i\varphi)]\right\}p(\varphi)\right]^2
        w(\varphi) d\varphi =
\end{equation*}
\begin{equation}\label{lambda}
    \int_{-\pi}^\pi \left[g(\varphi) - p(\varphi) +
        p(\varphi)\sum_{i=1}^4 C_i[2 - 2\cos(m_i\varphi)]\right]^2 w(\varphi) d\varphi,
\end{equation}
where $w(\varphi) \approx 1$ is additional weight. As usual, calculating partial derivatives
\begin{equation*}
    0 = \frac12\frac{\partial\lambda}{\partial C_k} =
    \int_{-\pi}^\pi \left[g(\varphi) - p(\varphi) +
        p(\varphi)\sum_{i=1}^4 C_i[2 - 2\cos(m_i\varphi)]\right]
        p(\varphi)w(\varphi)[2 - 2\cos(m_k\varphi)] d\varphi =
\end{equation*}
\begin{equation}
    \int_{-\pi}^\pi [g(\varphi) - p(\varphi)]p(\varphi)w(\varphi)[2 - 2\cos(m_k\varphi)] d\varphi +
    \sum_{i=1}^4 C_i\int_{-\pi}^\pi p^2(\varphi)w(\varphi)
        [2 - 2\cos(m_i\varphi)][2 - 2\cos(m_k\varphi)] d\varphi.
\end{equation}
After defining
\begin{equation}
    U_k \equiv \frac1{2\pi}\int_{-\pi}^\pi p^2(\varphi)w(\varphi)e^{-ip\varphi}d\varphi
\end{equation}
and
\begin{equation}
    V_k \equiv \frac1{2\pi}\int_{-\pi}^\pi g(\varphi)p(\varphi)w(\varphi)e^{-ip\varphi}d\varphi
\end{equation}
we can get
\begin{equation}
    0 = \frac1{4\pi}\frac{\partial\lambda}{\partial C_k} = 2[V_0 - V_{m_k} - U_0 + U_{m_k}] +
        2\sum_{i=1}^4 C_i[2U_0 - 2U_{m_i} - 2U_{m_k} + U_{m_i+m_k} + U_{|m_i-m_k|}].
\end{equation}
Resulting symmetric system of four linear equations
\begin{equation}
    \sum_{i=1}^4 C_i[2U_0 - 2U_{m_i} - 2U_{m_k} + U_{m_i+m_k} + U_{|m_i-m_k|}] =
        U_0 - U_{m_k} - V_0 + V_{m_k},
\end{equation}
can be solved through standard methods of linear algebra.

The real problem here is to efficiently calculate $U_k$ and $V_k$. One method is to use crude
approximation
\begin{equation}
    p(\varphi) \approx \exp\left\{-\frac{\alpha\varphi^2}2\right\},
\end{equation}
where
\begin{equation}\label{alpha}
    \alpha = \frac l2 + \frac56(1 - 4^{-n}).
\end{equation}
In that case
\begin{equation}
    U_k \approx \frac1{2\pi}\int_{-\pi}^\pi
        \exp\left\{-\alpha\varphi^2\right\}e^{-ip\varphi}d\varphi \approx \Gamma(2\alpha;\, k),
\end{equation}
\begin{equation}\label{V-gauss}
    V_k \approx \frac1{2\pi}\int_{-\pi}^\pi
        \exp\left\{-\left[\frac{\sigma^2}{4^n} + \alpha\right]\frac{\varphi^2}2\right\}
        e^{-ip\varphi}d\varphi \approx \Gamma(\sigma^24^{-n} + \alpha;\, k),
\end{equation}
where $\Gamma$ is normalized Gaussian function defined in (\ref{gamma}).

Much better method is to choose $w(\varphi)$ such that
\begin{equation}
    p(\varphi)w(\varphi) = \exp\left\{-\frac{\alpha\varphi^2}2\right\}
\end{equation}
holds exactly. In that case (\ref{V-gauss}) holds exactly too and for $U_k$ we
have
\begin{equation}
    U_k = \frac1{2\pi}\int_{-\pi}^\pi \exp\left\{-\frac{\alpha\varphi^2}2\right\}
        p(\varphi)e^{-ip\varphi}d\varphi =
    \frac1{2\pi}\int_{-\pi}^\pi \exp\left\{-\frac{\alpha\varphi^2}2\right\}
        B^{2l}(\varphi)D(\varphi)e^{-ip\varphi}d\varphi,
\end{equation}
where $D$ is effective rescale filter kernel with
\begin{equation}
    D(\varphi) \equiv \prod_{k=1}^n \cos^{10}\frac\varphi{2^{k+1}}.
\end{equation}
Therefore $U_k$ can be calculated in the following way: construct Gaussian function
$\Gamma(\alpha;\, k)$, then apply filter $\hat D$, than apply binomial filter of order $2l$.
Alternatively that can be written in operator notation as
\begin{equation}
    \hat U = \hat B^{2l}\hat D\,\hat\Gamma(\alpha;\, \cdot).
\end{equation}

@@

Third method is usually slightly better than second and much better then first, but precise methods
require good approximation of effective rescale filter $D$. Experiments shows that tight 7-tap
approximation
\begin{equation}
    D(\varphi) \approx D_0 + 2\sum_{k=1}^3 D_k\cos(k\varphi)
\end{equation}
is quite good. Coefficients here should be chosen in such a way that equation is exact up to 6
term of Taylor series around $\varphi = 0$. Corresponding problem can be solved analytically
and give the following results
\newcommand\0{\phantom{0}}
\begin{align}
    D_0 &= (   5204 +   2520\cdot4^{-n} +   1092\cdot16^{-n} +   3280\cdot64^{-n}) / 12096,\\
    D_1 &= (   2943 - \0 210\cdot4^{-n} - \0 273\cdot16^{-n} -   2460\cdot64^{-n}) / 12096,\\
    D_2 &= (\0  486 - \0 924\cdot4^{-n} - \0 546\cdot16^{-n} + \0 984\cdot64^{-n}) / 12096,\\
    D_3 &= (\0\0 17 - \0 126\cdot4^{-n} + \0 273\cdot16^{-n} - \0 164\cdot64^{-n}) / 12096.
\end{align}
Fig.~\ref{D} shows that such approximation noticeably differ from exact curve only at region near
$\varphi = \pm\pi$ where the target Gaussian is near zero anyway. Accurate approximation of
effective rescale filter kernel concludes calculation of the main filter coefficients, and in the
following section I study the task of choosing remaining parametrs $n$, $l$ and $m_i$.

@@

Combined effect of all stages of Gaussian blur approximation can be written in generic form
\begin{equation}
    R_i = \sum_{j\in Z} K_{ij}F_j,
\end{equation}
where
\begin{equation}
    K_{ij} \approx G_{i-j} = \Gamma(\sigma^2;\, i - j) =
        \frac1{\sqrt{2\pi\sigma^2}}\exp\left\{-\frac{(i - j)^2}{2\sigma^2}\right\}.
\end{equation}
Assuming that filter doesn't distort constant function (if $F_i = c$ then $R_i = c$) or
alternatively
\begin{equation}
    \sum_{j\in Z} K_{ij} = 1,
\end{equation}
and source values lie between $0$ and $1$,
\begin{equation}
    0 \le F_i \le 1,
\end{equation}
we can calculate exact upper bound of error $E^{\max}_i = \sup_F |E_i|$, where
\begin{equation}\label{E}
    E_i = \sum_{j\in Z} [K_{ij} - G_{i-j}]F_j.
\end{equation}
Due to zero error in case of zero source function
\begin{equation}
    E^{\max}_i = \sup_F |E_i| = \max\{-\inf_F E_i,\; \sup_F E_i\}.
\end{equation}
To calculate maximum positive and negative error one can choose $F_j$ in such a way that sum from
(\ref{E}) reduces to sum of only positive or negative terms,
\begin{align}
    \inf_F E_i &= \sum_{j\in Z, K_{ij} < G_{i-j}} [K_{ij} - G_{i-j}],\\
    \sup_F E_i &= \sum_{j\in Z, K_{ij} > G_{i-j}} [K_{ij} - G_{i-j}].
\end{align}
It's easy to see that
\begin{align}
    \sup_F E_i + \inf_F E_i &= \sum_{j\in Z} [K_{ij} - G_{i-j}] = 1 - 1 = 0,\\
    \sup_F E_i - \inf_F E_i &= \sum_{j\in Z} |K_{ij} - G_{i-j}|,
\end{align}
and
\begin{equation}
    E^{\max}_i = \max\{-\inf_F E_i,\; \sup_F E_i\} = \sup_F E_i =
        \frac12\sum_{j\in Z} |K_{ij} - G_{i-j}|.
\end{equation}
That's the final form of error estimation I'll use later.
